---
title: "Homework_2"
author: "Stromieri, Cusma Fait, Esposito, Lucas"
date: "2023-10-27"
output: html_document
---

```{r, echo=FALSE}
set.seed(0)
```

## FSDS - Chapter 3



### Ex 3.12

Simulate random sampling from a normal population distribution with several n values to
illustrate the law of large numbers.


**Solution**

We simulate a list of independent samples from a normal population.

```{r 3_12}
n <- 100
x <- rnorm(n)
y <- rep(0, n)

for (i in 1:n) {
  y[i] <- mean(x[1:i])
}

x_axis = 1:n
plot(x_axis, abs(y), type='l', ylim=c(-0.1, +1.1), 
     main='Absolute sample mean', xlab='n', ylab='abs mean')
lines(c(1, n), c(0, 0), col='red')
lines(x_axis, +x_axis^(-1/2), col='red', lty='dashed')

```

The solid black line shows how the absolute sample mean is converging to 0 as
the sample size increases, while the dashed red line is its expected behavior. 



### Ex 3.18

Sunshine City, which attracts primarily retired people, has 90,000 residents with a mean age
of 72 years and a standard deviation of 12 years. The age distribution is skewed to the left. A
random sample of 100 residents of Sunshine City has $y = 70$ and $s = 11$.

(a) Describe the center and spread of the (i) population distribution, (ii) sample data distribution. What shape does the sample data distribution probably have? Why?

(b) Find the center and spread of the sampling distribution of $Y$ for $n = 100$. What shape
does it have and what does it describe?

(c) Explain why it would not be unusual to sample a person of age 60 in Sunshine City, but
it would be highly unusual for the sample mean to be 60, for a random sample of 100
residents.

(d) Describe the sampling distribution of $Y$ : (i) for a random sample of size n = 1; (ii) if you
sample all 90,000 residents.


**Solution**

...

```{r 3_18}
```

...



### Ex 3.28

A survey is planned to estimate the population proportion π supporting more government
action to address global warming. For a simple random sample, if $\pi$ may be near 0.50, how
large should n be so that the standard error of the sample proportion is 0.04?

**Solution**

The problem suggests to start from an initial uncertainty of a binomial distribution with parameter $p=1/2$, which is $\sigma_0 = 1/2$ 

$$SE = \frac{\sigma_0}{\sqrt n} $$
$$n =  \frac{\sigma_0^2}{SE^2} = \frac{0.5^2}{0.04^2} = 156.25 $$

In order for the standard error of the sample proportion to be $SE=0.04$, the
sample size should be at least around 156.



## FSDS - Chapter 4



### Ex 4.2

For a sequence of observations of a binary random variable, you observe the geometric random
variable (Section 2.2.2) outcome of the first success on observation number $y = 3$. Find and plot
the likelihood function.


**Solution**

...

```{r 4_2}
```

...



### Ex 4.4

For the ```Students``` data file (Exercise 1.2 in Chapter 1) and corresponding population, find the
ML estimate of the population proportion believing in life after death. Construct a Wald 95%
confidence interval, using its formula (4.8). Interpret.

**Solution**

...

```{r 4_4}
```

...



### Ex 4.38

For independent observations $y_1, . . . , y_n$ having the geometric distribution (2.1):

(a) Find a sufficient statistic for $\pi$.

(b) Derive the ML estimator of $\pi$.

**Solution**

...

```{r 4_38}
```

...



### Ex 4.44

Refer to the previous two exercises. Consider the selling prices (in thousands of dollars) in the
Houses data file mentioned in Exercise 4.31.

(a) Fit the normal distribution to the data by finding the ML estimates of µ and σ for that
distribution.

(b) Fit the log-normal distribution to the data by finding the ML estimates of its parameters.

(c) Find and compare the ML estimates of the mean and standard deviation of selling price
for the two distributions.

(d) Superimpose the fitted normal and log-normal distributions on a histogram of the data.
Which distribution seems to be more appropriate for summarizing the selling prices?


**Solution**

...

```{r 4_44}
```

...



### Ex 4.48

For a simple random sample of n subjects, explain why it is about 95% likely that the sample
proportion has error no more than $1/\sqrt n$ in estimating the population proportion. (Hint: To
show this “$1/\sqrt n$ rule,” find two standard errors when $\pi = 0.50$, and explain how this compares to two standard errors at other values of $\pi$.) Using this result, show that $n = 1/M^2$
is a safe sample size for estimating a proportion to within M with 95% confidence.

**Solution**

...

```{r 4_48}
```

...



## FSDS - Chapter 5



### Ex 5.2

When a government does not have enough money to pay for the services that it provides, it
can raise taxes or it can reduce services. When the Florida Poll asked a random sample of 1200
Floridians which they preferred, 52% (624 of the 1200) chose raise taxes and 48% chose reduce
services. Let π denote the population proportion of Floridians who would choose raising taxes.
Analyze whether this is a minority of the population ($\pi < 0.50$) or a majority ($\pi > 0.50$) by
testing $H_0 ∶ \pi = 0.50$ against $H_a: \pi \ne 0.50$. Interpret the P-value. Is it appropriate to “accept $H_0$? Why or why not?


**Solution**

...

```{r 5_2}
```

...



### Ex 5.12

The example in Section 3.1.4 described an experiment to estimate the mean sales with a
proposed menu for a new restaurant. In a revised experiment to compare two menus, on Tuesday
of the opening week the owner gives customers menu A and on Wednesday she gives them menu
B. The bills average \$22.30 for the 43 customers on Tuesday ($s = 6.88$) and \$25.91 for the 50
customers on Wednesday ($s = 8.01$). Under the strong assumption that her customers each night
are comparable to a random sample from the conceptual population of potential customers,
show how to compare the mean sales for the two menus based on (a) the P-value of a significance
test, (b) a 95% confidence interval. Which is more informative, and why? (When used in an
experiment to compare two treatments to determine which works better, a two-sample test is
often called an A/B test.) 

**Solution**

...

```{r 5_12}
```

...



### Ex 5.68

Explain why the confidence interval based on the Wald test of $H_0: \theta = \theta_0$ is symmetric around $\hat \theta$ (i.e., having center exactly equal to $\hat \theta$. This is not true for the confidence intervals based on the likelihood-ratio and score tests.) Explain why such symmetry can be problematic when $\theta$ and $\hat \theta$ are near a boundary, using the example of a population proportion that is very close to 0 or 1 and a sample proportion that may well equal 0 or 1.

**Solution**

...

```{r 5_68}
```

...

